---
title: "30DCC_22-10"
author: "Christian Gebhard"
date: "2022-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## prepare data

```{r}
library(tidyverse)
library(ggtext)
library(exoplanets)
library(geomtextpath)
library(ggdark)

library(showtext)
font_add_google("Jura")
font_add_google("Audiowide")
font_add_google("Open Sans")
showtext_auto()
```

## access / load data

```{r, eval=FALSE}
# planet_table <- exoplanets("ps", columns = c("pl_name", "hostname", "pl_orbsmax"))
# write_csv(planet_table, "assets/planetary_data.csv")
```

```{r}
planets <- read_csv("assets/planetary_data.csv") %>% 
  select(pl_name, hostname, pl_orbsmax)

```


## create plot
```{r, layout="l-page", fig.width=8}
# prepare orbit data to requirements:

planet_orbit <- planets %>% 
  # calculate median of multiple measurements (available for some planets)
  group_by(pl_name) %>% 
  summarise(mean_major_axis = median(pl_orbsmax, na.rm = TRUE)) %>% 
  # drop missing values and filter out outliers (not many, still keeps ~94% of data points)
  drop_na(mean_major_axis) %>% 
  filter(mean_major_axis < 100) |> 
  # randomly scatter/jitter the distance values on a hypothetical x-axis
  mutate(x = runif(min = 0, max = 1, n = 4711))



annotation_data1 <- tribble(
  ~x, ~y, 
  0, 0.01,
  1, 0.01
)
annotation_data2 <- tribble(
  ~x, ~y, 
  0, 1,
  1, 1
)

# final plot:
po <-planet_orbit %>% 
  ggplot(aes(x=x, y=mean_major_axis)) +
  scale_y_continuous(
    trans = "log10",
    limits = c(0.00001, 100)
  ) +
  coord_polar(theta = "x") +
  labs(
    title = "<span><b>Inner Circle: Where Exoplanets Live</b></span>",
    subtitle = "<span>The number of discovered Exoplanets has passed the 5000 mark in<br>March 2022. Most of them are closer to their host star(s) than earth is<br>to the sun (~1 AU). This <b>experimental plot</b> shows the log10-scaled<br>distance distribution mapped to the y-Axis on polar coordinates<br>as a semi-quantitative 2D-density plot.</span>",
    caption = "<span>Data by NASA Exoplanet Archive, doi:	10.26133/NEA12<br>DataViz by @c_gebhard | <b>#30DayChartChallenge, Day 10</b></span>") +
  guides(fill = guide_legend(byrow = FALSE, title = "<b>Density</b>"))+
  geom_textline(
    data = annotation_data1,
    aes(x=x, y=y),
    label = "<span style='font-family:Jura;color:#FFFFFF;font-size: 20pt;'><b>0.01 AU</b></span>",
    rich = TRUE,
    color = "#FFFFFF",
    linetype = "dashed",
    vjust = 0.25
  ) +
  geom_textline(
    data = annotation_data2,
    aes(x=x, y=y),
    label = "<span style='font-family:Jura;color:#FFFFFF;font-size: 20pt;'><b>1 AU</b></span>",
    rich = TRUE,
    color = "#FFFFFF",
    linetype = "dashed",
    vjust = 0.25
  ) +
  stat_density_2d_filled(alpha = 0.7) +
  scale_fill_viridis_d(
    option = "E",
    labels = c("<b>low</b>", "", "", "", "", "", "", "", "", "", "<b>high</b>")
    )+
  dark_mode() +
  theme(
    legend.position = "right",
    plot.title.position = "panel",
    legend.title = element_markdown(family = "Audiowide", size = 24),
    legend.text = element_markdown(family = "Jura", size = 20),
    axis.line.y = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.title = element_markdown(family = "Audiowide", size = 54),
    plot.subtitle = element_markdown(family = "Jura", size = 32, lineheight = 0.35),
    plot.caption = element_markdown(family = "Open Sans", size = 20, lineheight = 0.4),
    panel.background = element_blank(),
    plot.caption.position = "plot"
  ) 

# save to disk
ggsave(here::here("..", "plots", "2022_10.png"), height = 5.5, width = 5, dpi = "retina")


