---
title: "30DCC_22-08"
author: "Christian Gebhard"
date: "2022-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# data source:

Tobias et al 2022 Ecology Letters
https://doi.org/10.6084/m9.figshare.16586228.v5

## prepare data

```{r}
library(tidyverse)
library(ggtext)
library(MetBrewer)
library(patchwork)
library(ggdist)

library(showtext)
font_add_google("Playfair Display")
font_add_google("Lato")
showtext_auto()
```


## AVONET

The downloadable data archive contains several data sets. For this plot, the Supplementary dataset No. 1 was used. Apparently there are different species definitions. Without knowledge of the scientific details, I chose the 'Bird Tree' system, that was also used in many parts of the above mentioned paper.

```{r}
library(readxl)
AVONET_Supplementary_dataset_1 <- read_excel("~/Projekte/2022_30daysofcode/data/avonet/AVONET Supplementary dataset 1.xlsx", 
    sheet = "AVONET3_BirdTree") 

avonet3 <- AVONET_Supplementary_dataset_1|> 
  janitor::clean_names() |> 
  # reclassify missing habitat to "not categorized
  mutate(habitat = ifelse(habitat == "NA", "not categorized", habitat)) |> 
  # calculate grouped mean and median
  group_by(habitat) |> 
  mutate(
    tail_median = median(tail_length, na.rm = TRUE),
    tail_mean = mean(tail_length, na.rm = TRUE)
  ) |> 
  ungroup()

# bring to long form
avolong <- avonet3 |> 
  pivot_longer(cols = c(beak_length_nares, tarsus_length, wing_length, tail_length, mass), names_to = "long_measure", values_to = "long_value")


```   



# plot the median-sorted plot

```{r}
med_plot <- ggplot(avonet3, aes(x = reorder(habitat, tail_median), y = tail_length, color = reorder(habitat, tail_median), fill = reorder(habitat, tail_median))) +
  geom_point(
    size = 0.5,
    alpha = 0.1,
    position = position_jitter(
      seed = 1, width = .1
    ),
    shape = 20
  ) +
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.3, 
    point_colour = NA,
    alpha = 0.6) + 
  geom_point(
    aes(x = habitat, y = tail_mean),
    color = "#c62320",
    size = 6,
    shape = "|"
  ) +
  geom_point(
    aes(x = habitat, y = tail_median),
    color = "#0000FF",
    size = 6,
    shape = "|"
  ) +
  labs(
    title = "<b>Sorted by <span style='color: #0000FF'>median</span> (descending)</b>",
    y = "Tail length [mm]"
  ) +
  coord_flip() +
  scale_color_manual(values = met.brewer("Hokusai3", n = 12, type = "continuous")) +
  scale_fill_manual(values = met.brewer("Hokusai3", n = 12, type = "continuous")) +
  theme_classic()+
  theme(
    legend.position = "none",
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major.x = element_line(colour = "#aaaaaa", size = 0.1),
    plot.title.position = "plot",
    plot.title = element_markdown(family = "Lato", size = 20),
    plot.subtitle = element_markdown(family = "Lato", size = 32, lineheight = 0.4),
    axis.title.y = element_blank(),
    axis.title.x = element_markdown(family = "Lato", size = 20),
    axis.text = element_markdown(family = "Lato", size = 20),
    axis.text.y = element_markdown(face = "bold"),
    plot.caption = element_markdown(family = "Lato", size = 20, lineheight = 0.4)
  )




```

## plot the mean-sorted plot

```{r}
mean_plot <- ggplot(avonet3, aes(x = reorder(habitat, tail_mean), y = tail_length, color = reorder(habitat, tail_median), fill = reorder(habitat, tail_median))) +
  geom_point(
    size = 0.5,
    alpha = 0.1,
    position = position_jitter(
      seed = 1, width = .1
    ),
    shape = 20
  ) +
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.3, 
    point_colour = NA,
    alpha = 0.7) + 
  geom_point(
    aes(x = habitat, y = tail_mean),
    color = "#c62320",
    size = 6,
    shape = "|"
  ) +
  geom_point(
    aes(x = habitat, y = tail_median),
    color = "#0000FF",
    size = 6,
    shape = "|"
  ) +
  labs(
    title = "<b>Sorted by <span style='color: #c62320'>mean</span> (descending)</b>",
    y = "Tail length [mm]"
  ) +
  coord_flip() +
  scale_color_manual(values = met.brewer("Hokusai3", n = 12, type = "continuous")) +
  scale_fill_manual(values = met.brewer("Hokusai3", n = 12, type = "continuous")) +
  theme_classic()+
  theme(
    legend.position = "none",
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major.x = element_line(colour = "#aaaaaa", size = 0.1),
    plot.title.position = "plot",
    plot.title = element_markdown(family = "Lato", size = 20),
    plot.subtitle = element_markdown(family = "Lato", size = 32, lineheight = 0.4),
    axis.title.y = element_blank(),
    axis.title.x = element_markdown(family = "Lato", size = 20),
    axis.text = element_markdown(family = "Lato", size = 20),
    axis.text.y = element_markdown(face = "bold"),
    plot.caption = element_markdown(family = "Lato", size = 20, lineheight = 0.4)
  )


```


## compose final plot

```{r}

# setup a "background plot"
back_plot <- ggplot() +
  labs(
    title = "You're so mean!",
    subtitle = "When examining <b>group differences</b>, comparing means is not always a good option.<br>Measurements can be skewed, with considerably different <b>means and medians</b>.",
    caption = "<span><br>Data: 'AVONET' by Tobias et. al. 2022, doi:10.1111/ele.13898</span><br><span>DataViz by @c_gebhard | <b>#30DayChartChallenge, Day 09</b></span>"
  ) +
  theme_classic()+
  theme(
    legend.position = "none",
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major.x = element_line(colour = "#aaaaaa", size = 0.1),
    plot.title.position = "plot",
    plot.title = element_markdown(family = "Playfair Display", size = 72),
    plot.subtitle = element_markdown(family = "Lato", size = 32, lineheight = 0.4),
    axis.title.y = element_blank(),
    axis.title.x = element_markdown(family = "Lato", size = 20),
    axis.text = element_markdown(family = "Lato", size = 20),
    axis.text.y = element_markdown(face = "bold"),
    plot.caption = element_markdown(family = "Lato", size = 20, lineheight = 0.4)
  )
  
# position the mean- and median-sorted plots on the background
patched <- back_plot +
  inset_element(med_plot, left = 0, right = 0.5, top = 1, bottom = 0)+
  inset_element(mean_plot, left = 0.5, right = 1, top = 1, bottom = 0)

ggsave(here::here("..", "plots", "2022_09.png"), plot = patched, height = 5.5, width = 6, dpi = "retina")


```

